---
title: "Programming R Final"
author: "Matthias Rathbun"
date: "2022-10-30"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width = 20,fig.height = 20)
```


# Set up the work space

First I want to reduce the number of displayed digits.

```{r}
options(
  digits = 4
)
```

# Import Data

Import dataset as a dataframe

```{r}
Wine <- read.csv(
  file = "winequality-red.csv",
  header = TRUE
)
```

## Check Dataset Structure for tidyness

```{r}
head(Wine[,-1],10)
```
 Data is Clean and Tidy, each column is a feature, each row is a unique observation, and every cell has a single value.
 
# Visualize Data

## Corr Plot

I would Like to See the correlations between features as this is a regression problem
```{r}
library(GGally)
ggcorr(Wine, label = TRUE, label_round = 3)
ggsave("Corrplot.pdf")
```

## Pairplot

```{r}
ggpairs(Wine)
ggsave("pairplot.pdf")
```

Usefull for seing distributions and for correlations between variables.

# Scatter Plot with Loess

```{r}
ggplot(Wine, aes(fixed.acidity, density)) +
  geom_point(aes(colour = quality)) +
  geom_smooth(se = FALSE) +
  labs(
    x = "Acidity",
    y = "Density",
    colour = "Wine Quality"
  )
ggsave("scatter.pdf")
```
This shows some trend between Acidity and density of wine. These were chosen as they already had high correlation based on the corrplot

# Principle Component Analysis

```{r}
library(ggfortify)
pca_res <- prcomp(Wine, scale. = TRUE)

autoplot(pca_res, data = Wine , colour = 'quality', size = 4)

ggsave("pca.pdf")
```
Reduce 11 Features to 2 axes the higher quality wines seem to be at the bottom left part of the plot. Lower quality wines are clustered together.
